---
title: "STAT 206 Homework 6"
author: "Emilia Avanes"
output:
  word_document: default
  pdf_document: default
---

Due Tuesday, November 12, 5:00 PM

General instructions for homework: Homework must be submitted as pdf file, and be sure to include
your name in the file. Give the commands to answer each question in its own code block, which will also
produce plots that will be automatically embedded in the output file. Each answer must be supported by
written statements as well as any code used. (Examining your various objects in the “Environment” section
of RStudio is insufficient – you must use scripted commands.)

Part I - Gambler’s Ruin
============

```{r}
# 1. Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome
# directly for at most 5000 iterations of the game (or hands), estimate the following. (You must stop
# playing if your player has gone bust.)
# 
# a. the probability that you have “busted” (lost all your money) by the time you have placed your
# one hundredth bet.
# Probability of going bust by 100th hand  = 0.3156

# We simulate 5000 iterations of game where each game stops if the player has gone bust or has played 5000 hands without going bust.

bankroll = 1000
each_bet = 100

iterations = 5000

# probability to win a game. Lets assume there 50% chance of winning any game
p = 0.5
game_id = c()
hands_to_bust = c()
set.seed(1012)
all_bankroll = c()
for(i in 1:iterations){
  bankroll = 1000
  num_hands = 0
  while((bankroll > 0) & (num_hands < 5000)){ 
    # continue playing until you've gone bust. and number of hands is less than 5000.
    result = sample( c(0,1), 1, prob = c(1-p,p))
    bankroll = bankroll - each_bet + (result * 2 * each_bet)
    num_hands = num_hands + 1
  }
  game_id = c(game_id, i)
  hands_to_bust = c(hands_to_bust, num_hands)
  all_bankroll = c(all_bankroll, bankroll)
}

# probability of going bust by 100th hand  = 0.3156
sum(hands_to_bust<=100)/length(hands_to_bust)

# b. the probability that you have busted by the time you have placed your five hundredth bet by
# simulating the outcome directly.
# Probability of going bust by 500th hand  = 0.6472
sum(hands_to_bust<=500)/length(hands_to_bust)

# c. the mean time you go bust, given that you go bust within the first 5000 hands.
# Mean time to go bust, if you go bust within 5000 hands = 1053.022 hands.

req_hands = hands_to_bust[hands_to_bust<=5000] #select hands where we go bust in first 5000 hands.
mean(req_hands)


# d. the mean and variance of your bankroll after 100 hands (including busts).
# mean = $ 984
# variance = $ 9,792,065
# mean and variance of first 100 hands. 
req_bankroll = all_bankroll[1:100]
mean(req_bankroll)
var(req_bankroll)

# e. the mean and variance of your bankroll after 500 hands (including busts).
# mean = $ 1184
# variance = $ 11,858,822
# mean and variance of first 100 hands. 
req_bankroll = all_bankroll[1:500]
mean(req_bankroll)
var(req_bankroll)

# 2. Repeat the previous problem with betting on black in American roulette, where the probability of
# winning on any spin is 18/38 for an even payout.
# Everything remains the same except probability of winning .
p = 18/38
# We simulate 5000 iterations of game where each game stops if the player has gone bust or has played 5000 hands without going bust.

bankroll = 1000
each_bet = 100

iterations = 5000

# probability to win a game. Lets assume there 50% chance of winning any game
game_id = c()
hands_to_bust = c()
set.seed(1012)
all_bankroll = c()
for(i in 1:iterations){
  bankroll = 1000
  num_hands = 0
  while((bankroll > 0) & (num_hands < 5000)){ 
    # continue playing until you've gone bust. and number of hands is less than 5000.
    result = sample( c(0,1), 1, prob = c(1-p,p))
    bankroll = bankroll - each_bet + (result * 2 * each_bet)
    num_hands = num_hands + 1
  }
  game_id = c(game_id, i)
  hands_to_bust = c(hands_to_bust, num_hands)
  all_bankroll = c(all_bankroll, bankroll)
}

# a. the probability that you have “busted” (lost all your money) by the time you have placed your
# one hundredth bet.
# probability of going bust by 100th hand  = 0.4976

sum(hands_to_bust<=100)/length(hands_to_bust)

# b. the probability that you have busted by the time you have placed your five hundredth bet by
# simulating the outcome directly.
# probability of going bust by 500th hand  = 0.9216
sum(hands_to_bust<=500)/length(hands_to_bust)


# c. the mean time you go bust, given that you go bust within the first 5000 hands.
# mean time to go bust, if you go bust within 5000 hands = 186.5932 hands.

req_hands = hands_to_bust[hands_to_bust<=5000] #select hands where we go bust in first 5000 hands.
mean(req_hands)


# d. the mean and variance of your bankroll after 100 hands (including busts).
# mean = $ 0
# variance = $ 0
# mean and variance of first 100 hands. 
req_bankroll = all_bankroll[1:100]
mean(req_bankroll)
var(req_bankroll)

# e. the mean and variance of your bankroll after 500 hands (including busts).
# mean = $ 0
# variance = $ 0
# mean and variance of first 100 hands. 
req_bankroll = all_bankroll[1:500]
mean(req_bankroll)
var(req_bankroll)


# 3. For the American roulette problem in the previous question, you calculated a mean value. Because you
# saved these final results in a vector, use the bootstrap to estimate the variance of the return in each
# case for your final answer.

var(all_bankroll)
```

Part II - Elo Ratings
=========

One of the earliest examples of a convergent, adaptive Markov process was the rating system devised by
Arpad Elo to rank chess players. It has endured for so long as a simple system for so long that it is used
as a primary ranking system in many other scenarios, including the NBA team rankings (Nate Silver) and
Scrabble (NASPA).
The main idea is two players have ratings RA and RB. The estimated probability that player A will win is
modeled by a logistic curve,
P(A) = 1
1 + exp(RB − RA)
and once a game is finished, a player’s rating is updated based on whether they won the game:
RA(new) = RA(old) + K(1 − P(A))
or if the lost the game:
RA(new) = RA(old) − KP(A)
for some factor K. (Note that both player ratings change.) Our goal is to simulate a repetitive tournament
with 10,000 games to see if it converges on the true values.

```{r}
# 4. Create a “true” vector of ratings for 13 players whose ratings range from -2 to 2 in even intervals.
# Create another vector with the current ratings which will be updated on a game-by-game basis, and a
# matrix with 13 rows and 10,000 columns into which we will deposit the ratings over time.

set.seed(10121)

true_rat = seq(-2,2,length.out = 13)
curr_rat = 1:13#rep(,13) # assuming all have same current rating = 0
mat_rat = matrix(0,nrow = 13, ncol = 10000)

# 5. Write a function that simulates a game between players i and j given their true underlying ratings. This
# should be a simple draw from rbinom(1,1,p) with the appropriate probability.

win_prob = function(i, j){
  
  rating_i = true_rat[i]
  rating_j = true_rat[j]
  
  pA = 1/(1 + (exp(rating_j - rating_i)))
  return(pA)
}

play_game = function(i,j){

  pA = win_prob(i,j)
  rbinom(1,1,pA)
  
}

play_game(3,4)

# 6. Write a function that, given a value of K, replaces the ratings for the two players who just played a
# game with their updated ratings given the result from the previous question.

# i = sel_ply[1]
# j = sel_ply[2]

update_rating = function(i,j,K,result,obs_rating){
  
  # result = play_game(i,j)
  pA = win_prob(i,j)
  
  if(result == 1){ # player i wins
    
    obs_rating[i] = obs_rating[i] + K*(1-pA)
    obs_rating[j] = obs_rating[j] - K*(pA)
    
  } else {
    
    obs_rating[j] = obs_rating[j] + K*(1-pA)
    obs_rating[i] = obs_rating[i] - K*(pA)
    
    }
  obs_rating
}
  

# 7. Write a function that selects two players at random from the 13, makes them play a game according to
# their true ratings, and updates their observed ratings.
stimulate_a_game = function(K, obs_rating){
  
  sel_ply = sample(1:13,2)
  result = play_game(sel_ply[1],sel_ply[2])
  update_rating(sel_ply[1],sel_ply[2], K, result,obs_rating)
  
  }

# 8. Finally, write a function that simulates a tournament as prescribed above: 10,000 games should be
# played between randomly chosen opponents, and the updated ratings should be saved in your rating
# matrix by iteration.
# K = 0.01
stimulate_tournament = function(K){
  
  rating_now = rep(0, 13) # assuming all have same current rating = 0
  
  for(i in 1:10000){
    
    rating_now = stimulate_a_game(K, obs_rating = rating_now)
    # print(rating_now)
    mat_rat[,i] <<- rating_now
    
  }
}

# 9. Run this tournament with K = 0.01. Plot the rating for the best player over time using plot(...,
# ty="l"); add the rating for the worst player using lines(...). Do they appear to converge to the
# true ratings?

stimulate_tournament(K= 0.01)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')

# 10. Repeat the previous step with K equal to 0.03, 0.06, 0.1, 0.3, 0.6 and 1. Which appears to give the
# most reliable rating results?
stimulate_tournament(K= 0.03)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')

stimulate_tournament(K= 0.06)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')

stimulate_tournament(K= 0.1)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')

stimulate_tournament(K= 0.3)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')

stimulate_tournament(K= 0.6)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')

stimulate_tournament(K= 1)
mat_rat[,10000]

plot(as.vector(mat_rat[13,]),ty = 'l')
plot(as.vector(mat_rat[1,]),ty = 'l')


```
