---
title: "STAT 206 Homework 2"
author: "Emilia Avanes"
date: "10/4/2019"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STAT 206 Homework 2

## Question 1:
```{r}

## 1a. Load the data into a dataframe called `ca_pa':

ca_pa.df = read.csv(file = 'http://www.stat.cmu.edu/~cshalizi/uADA/13/hw/01/calif_penn_2011.csv')

## 1b. How many rows and columns does the dataframe have? rows: 11275, columns:34
dim(ca_pa.df)

## 1c. This line sums up the number of times "na" appears in each column.
colSums(apply(ca_pa.df,c(1,2),is.na))

## 1d. The function `na.omit()` takes a dataframe and returns a new dataframe, omitting any row containing an NA value.  Use it to purge the data set of rows with incomplete data.
na.omit(ca_pa.df)

## 1e. Number of rows omitted since they have incomplete data.670 rows are omitted due to one of the columns having "na'
nrow(ca_pa.df)- nrow(na.omit(ca_pa.df))

## 1f. No, the values from c and e are not compatible. In c we are adding all the na's in each column, with the highest number of na's being found in the Median_house_value column (599). In e we are removing the rows with any na value (670). It makes sense that the second value will be greater than or equal to 599 since there is at least one column with that many na's.

```

## Question 2

```{r}

## 2a. The variable `Built_2005_or_later` indicates the percentage of houses in each Census tract built since 2005.  Plot median house prices against this variable.

median = (ca_pa.df[,10])
year = (ca_pa.df[,16])

plot(median, year, xlab = "Median house prices", ylab = "Built_2005_or_later")


## 2b. Make a new plot, or pair of plots, which breaks this out by state.  Note that the state is recorded in the `STATEFP` variable, with California being state 6 and Pennsylvania state 42.

median = (ca_pa.df[ca_pa.df[3]==6,10])
year = (ca_pa.df[ca_pa.df[3]==6,16])

plot(median, year, xlab = "Median house prices in CA", ylab = "Built_2005_or_later")

median = (ca_pa.df[ca_pa.df[3]==42,10])
year = (ca_pa.df[ca_pa.df[3]==42,16])

plot(median, year, xlab = "Median house prices in PA", ylab = "Built_2005_or_later")

```

## Question 3:

```{r}

## 3a.The vacancy rate is the fraction of housing units which are not occupied. The dataframe contains columns giving the total number of housing units for each Census tract, and the number of vacanthousing units. Add a new column to the dataframe which contains the vacancy rate. What are the minimum,maximum, mean, and median vacancy rates? Minimum is 0, Maximum is 0.965311, and mean is 0.08888789.

ca_pa.df$Vacancy_rate = (ca_pa.df[,12]/ca_pa.df[,11])

min (na.omit(ca_pa.df)$Vacancy_rate)
max (na.omit(ca_pa.df)$Vacancy_rate)
mean (na.omit(ca_pa.df)$Vacancy_rate)

## 3b. Plot the vacancy rate against median house value.

plot(ca_pa.df$Vacancy_rate, ca_pa.df$Median_house_value, xlab = "Vacancy Rate", ylab = "Median house prices")

## 3c. Plot vacancy rate against median house value separately for California and for Pennsylvania. Is there a difference? Yes comparing the total vacancy rate graph then looking separately at CA and PA it shows that in PA there is higher vacancy rates for houses with lower median house prices than in CA. 

CA_vacancy = (ca_pa.df[ca_pa.df[3]==6,"Vacancy_rate"])
median = (ca_pa.df[ca_pa.df[3]==6,10])

plot(CA_vacancy, median, xlab = "Vacancy Rate- CA", ylab = "Median house prices")

PA_vacancy = (ca_pa.df[ca_pa.df[3]==42,"Vacancy_rate"])
median = (ca_pa.df[ca_pa.df[3]==42,10])

plot(PA_vacancy, median, xlab = "Vacancy Rate- PA", ylab = "Median house prices")

```

## Question 4:

``` {r}
## The column COUNTYFP contains a numerical code for counties within each state. We are interested in Alameda County (county 1 in California), Santa Clara (county 85 in California), and Allegheny County (county 3 in Pennsylvania).
## 4a. Explain what the block of code at the end of this question is supposed to accomplish, and how it does it. 
## The block of code below reports the median value for the median house value in Alameda CA. It does this by first creating a vector for Alameda County in CA by checking if the StateFP is 6 and the CountyFP is 1 by going through the whole list one by one and appending each value to the vector. It then creates another vector for the Alameda County Median House Value (accamhv) which again goes through the prior vector and appends all the values in column for Median House Value (column 10) then it takes the function to get the median of the vector.

```{r, eval=FALSE}
ca_pd.df=na.omit(ca_pa.df)
acca <- c()
for (tract in 1:nrow(ca_pd.df)) {
  if (ca_pd.df$STATEFP[tract] == 6) {
    if (ca_pd.df$COUNTYFP[tract] == 1) {
      acca <- c(acca, tract)
    }
  }
}
accamhv <- c()
for (tract in acca) {
  accamhv <- c(accamhv, ca_pd.df[tract,10])
}
median(accamhv)


## 4b. Give a single line of R which gives the same final answer as the block of code. Note: there are at least two ways to do this; you just have to find one.


median(ca_pd.df[which(ca_pd.df$STATEFP == 6 & ca_pd.df$COUNTYFP == 1),10])


## 4c. For Alameda, Santa Clara and Allegheny Counties, what were the average percentages of housing built since 2005?
## Alameda
Alameda_mean= mean(ca_pd.df[which(ca_pd.df$STATEFP == 6 & ca_pd.df$COUNTYFP == 1),16])
print (Alameda_mean)

##Santa Clara
mean(ca_pd.df[which(ca_pd.df$STATEFP == 6 & ca_pd.df$COUNTYFP == 85),16])

## Allegheny Counties
mean(ca_pd.df[which(ca_pd.df$STATEFP == 6 & ca_pd.df$COUNTYFP == 3),16])


## 4d. The cor function calculates the correlation coefficient between two variables. What is the correlation between median house value and the percent of housing built since 2005 in (i) the whole data, (ii) all of California, (iii) all of Pennsylvania, (iv) Alameda County, (v) Santa Clara County and (vi) Allegheny County?

cor (ca_pd.df$Median_house_value, ca_pd.df$Built_2005_or_later)
for (ca_pd.df$STATEFP[tract] == 6) 
{
 corr= cor (ca_pd.df$Median_house_value, ca_pd.df$Built_2005_or_later)

}

## 4e. Make three plots, showing median house values against median income, for Alameda, Santa Clara,and Allegheny Counties. (If you can fit the information into one plot, clearly distinguishing the three counties, that’s OK too.)


```




